---
output:
  html_document: default
  pdf_document: default
  word_document: default
params:
  symbol: IBM
---

## `r params$symbol`

```{r, echo=FALSE, message=FALSE}

library(quantmod)

prices <- round(getSymbols(params$symbol, auto.assign = FALSE, src = 'google'), 2)
close <- Cl(last(prices))
open <- Op(last(prices))
```

The stock closed `r ifelse(open>close,'up','down')` at `r close` dollars per share yesterday ... hey this Phil 3/21/2017 Have a great day!!!

## Price History

The chart below is made with the `quantmod` R package, a widely used package for collecting and visualizing financial data in R. You can learn more about `quantmod` at the website this is my report
***
  
```{r echo=FALSE}
chartSeries(prices, theme = chartTheme("white", bg.col = "white"))
```

## Raw Data

The table below displays the daily price data for the stock. In the next section, we will learn how to make a concise, interactive table with the `DT` package, a new package for making searchable data tables. You can learn more about the `DT` package at the website 
***
  
```{r echo=FALSE}
DT::datatable(data.frame(prices[, 1:4], 2))
```

